### 2008-09-10 doc/a-rex/arex_tech_doc.tex
rename cluster+qju.pl to CEinfo.pl
Implementing comments from Gabor. Commentig LRMS specific parts.

\begin{tabular}{|c|c|c|>{\raggedright}p{10cm}|}
\hline 
Applicable BES State&
ARC Sub-state&
A-REX internal state&
\multicolumn{1}{c|}{Description}\tabularnewline
\hline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{2}{2cm}{Pending}&
Accepting&
ACCEPTED&
Job is in the process of being submitted. This state is not recognised
by the A-REX yet. \emph{Accepted} is first reported state.\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Accepted&
ACCEPTED&
Job was submitted\tabularnewline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{7}{2cm}{Running}&
Preparing&
PREPARING&
Stage-in process is going on\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Prepared&
PREPARING+PENDING&
Stage-in process has finished\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Submiting&
SUBMIT&
Communication with local batch system is in process\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
&
\multicolumn{1}{c|}{Queued}&
\multicolumn{1}{c|}{INLRMS}&
\multicolumn{1}{p{10cm}|}{Job entered local batch system but is not runnning now. This state
is not recognised by the A-REX yet. \emph{Executing} is reported instead }\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Executing&
INLRMS&
Job is being executed in local batch system\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Executed&
INLRMS,INLRMS+PENDING&
Job execution in local batch system has finished. The A-REX dos not
detect job states inside local batch system yet. As result this state
is reported only if job is \emph{Pending}.\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Killing&
CANCELING&
Communication with local batch system to terminate execution is in
process\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Finishing&
FINISHING&
Stage-out process is going on\tabularnewline
\hline 
Cancelled&
Killed&
FINISHED&
Job was stopped by explicit user request. The A-REX currently does
not remember this request. \emph{Failed} is reported instead.\tabularnewline
\hline 
\multicolumn{1}{|c|}{Failed}&
\multicolumn{1}{c|}{Failed}&
FINISHED&
\multicolumn{1}{l|}{There was a failure during execution}\tabularnewline
\hline 
Finished&
Finished&
FINISHED&
Job finished successfully\tabularnewline
\hline 
Finished&
Deleted&
DELETED&
Job finished and was left in A-REX too long\tabularnewline
\hline 
All&
Pending&
PENDING&
Job is prevented from going to the next state due to some internal
limits; this sub-state appears in parallel with other sub-states\tabularnewline
\hline 
All&
Held&
&
Job processing is suspended on client request; this sub-state appears
in parallel with other sub-states. This state is reserved for future
and is not implemented yet.\tabularnewline
\hline
\end{tabular}
\end{table}

@@

\begin{itemize}
\item \emph{gm-kick} - Sends a signal to the A-REX though a FIFO file to
wake it up. It's used to increase responsiveness of A-REX.
\item \emph{cluster+qju.pl} - Collects and generates information about computing
resource as XML document in NorduGrid and Glue 2 format.
\end{itemize}
The following modules are always run under the Unix account to which
a Grid user is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Currently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
pseudo-LRMS is supported for testing purposes. This module is responsible
for job submission to the LRMS.
\item \textit{cancel-{*}-job} - This script is for canceling jobs which
have been already submitted to the LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the A-REX about completion of jobs. It's implementation for PBS uses
server logs to extract information about jobs. If logs are not available
it uses the less reliable \emph{qstat} command for that. Other backends
use different techniques.
\end{itemize}
\begin{comment}In addition, there is also an administration utility:

@@

\begin{itemize}
\item \textit{session root directory} - This is the directory in which a
user's SDs are created. It's location is configurable per UNIX user.
Several (or even all) users may share the same session root directory.\\
The A-REX need to have permission to create new files and directories
in the session root directory. If A-REX is run under a dedicated user
account, that account needs full permissions in the \textit{session
root directory}.\\
If A-REX is run under the \emph{root} account, make sure \textit{session
root directory} resides on a file system which does not limit the
capabilities of the \emph{root} user (as does for example NFS with
\emph{root\_squash} option).\\
If there is a need to run A-REX under the \emph{root} account (to
be able to run jobs in LRMS under different users' accounts, for example)
but there is no way to provide a suitable \textit{session root directory,}
use the \emph{norootpower} command in configuration file. In that
case A-REX will use the identity of the local user to which a Grid
identity is mapped to access the \textit{session root directory}.
Hence those users will need full access there.\\
The A-REX creates SDs with proper ownership and permissions for the
local identity used to run a job. Some file systems require users
to have \emph{execute} permission on the \textit{session root directory}
in order to access any file or subdirectory there.\\
In order for jobs to access their input files, session root directories
should be shared across cluster nodes. Otherwise, LRMS-specific methods
must be used to transfer files to execution nodes. For more information
see section \ref{sub:LRMS}.
\item \textit{control directory} - In this directory A-REX stores an information
about accepted jobs. It must have full permissions there.\\
A subdirectory called \textit{log} is created there. It is used to
accumulate information about started and finished jobs. This information
is periodically sent to the desired \emph{logger service}(s). For
each job start and stop event, and for each logger service where that
event must be sent, a separate file is written. Once an event is sent,
the corresponding file is deleted.
\end{itemize}

@@

\begin{itemize}
\item \textbf{\textit{control}}\textit{=path username {[}username {[}...]]}
- This option initiates UNIX user as being serviced by the A-REX.
\textit{path} refers to the control directory (see section \ref{section: internals}
for the description of control directory). If the path is {*} the
default one is used - \$HOME/.jobstatus . \textit{username} stands
for UNIX name of the local user. Multiple names can be specified.
If the name is {*} it is substituted by all names found in file /etc/grid-security/grid-mapfile
(for the format of this file one should study the Globus project \cite{globus}).\\
Also the special name '.'(dot) can be used. Corresponding control
directory will be used for \textbf{any} user. This option should be
the last one in the configuration file. There is also command \textbf{\textit{controldir}}\textit{=path}.
It uses special username '.' and is always executed last independent
of placement in file.
\item \textbf{\textit{helper}}\textit{=username command {[}argument {[}argument
{[}...]]]} - associates external program with the local UNIX user.
This program will be kept running under account of the specified user.
\textit{username} stands for the name of the user. Special names can
be used: '{*}' - all names from /etc/grid-security/grid-mapfile, '.'
- root user. The user should be already configured with \textit{control}
option (except root, who is always configured). \textit{command} is
an executable and \textit{argument}s are passed as arguments to it.
\end{itemize}
Following are global commands specific to communication with underlying
LRMS (PBS in this case).

\begin{itemize}
\item \textbf{\textit{pbs\_bin\_path}}\textit{=path} - path to directory
which contains PBS commands.
\item \textbf{\textit{pbs\_log\_path}}\textit{=path} - path to directory
with PBS server's log files.
\item \textbf{\textit{gnu\_time}}\textit{=path} - path to \emph{time} utility.
\item \textbf{\textit{tmpdir}}\textit{=path} - path to directory for temporary
files.
\item \textbf{\textit{runtimedir}}\textit{=path} - path to directory which
contains \emph{runtimenvironment} scripts.
\item \textbf{\textit{shared\_filesystem}}\textit{=yes|no} - if computing
nodes have an access to session directory through a shared file system
like NFS. Corresponds to an environment variable RUNTIME\_NODE\_SEES\_FRONTEND
(see section \ref{sub:LRMS}).
\item \textbf{\textit{nodename}}\textit{=command} - command to obtain hostname
of computing node.
\item \textbf{\textit{scratchdir}}\textit{=path} - path on computing node
where to move session directory before execution.
\item \textbf{\textit{shared\_scratch}}\textit{=path} - path on frontend
where \textbf{\textit{scratchdir}} can be found.
\end{itemize}
In the command arguments (paths, executables, ...) following substitutions
can be used:

@@

@@

@@

Please read \emph{User Manual {[}??]} for usage instructions.

@@

\medskip{}
Provided by uploader component (URL is replaced by destination of
output file) :\\
\begin{longtable}{|p{2.5in}|p{4in}|}
\hline 
\emph{Error string}&
\emph{Reason/description}\tabularnewline
\hline 
Internal error in uploader&
Generic error \tabularnewline
\hline 
Output file: URL - unknown error&
Generic error \tabularnewline
\hline 
Output file: URL - unexpected error&
Generic error \tabularnewline
\hline 
User requested to store output locally URL&
Destination is URL of type \emph{file}.\tabularnewline
\hline 
Output file: URL - bad source URL&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - bad destination URL&
Destination URL is either malformed or not supported\tabularnewline
\hline 
Output file: URL - failed to resolve source locations&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed to resolve destination locations&
Problems related to Data Indexing service.\tabularnewline
\hline 
Output file: URL - failed to register new destination file&
-//-\tabularnewline
\hline 
Output file: URL - can't start reading from source&
User request to store output file, but there is no such file or there
are problems accessing session directory\tabularnewline
\hline 
Output file: URL - can't start writing to destination&
Problems with Data Storing services\tabularnewline
\hline 
Output file: URL - can't read from source&
Problems accessing session directory\tabularnewline
\hline 
Output file: URL - can't write to destination&
Problems with Data Storing services\tabularnewline
\hline 
Output file: URL - data transfer was too slow&
Timeout during transfer\tabularnewline
\hline 
Output file: URL - failed while closing connection to source&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed while closing connection to destination&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed to register new location&
Problems related to Data Indexing service.\tabularnewline
\hline 
Output file: URL - can't use local cache&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - system error&
Operating System returned error code where unexpected\tabularnewline
\hline 
Output file: URL - delegated credentials expired&
Access to destination requires credententials and they are either
outdated or missing (not delegated).\tabularnewline
\hline 
&
\tabularnewline
\hline
&
\tabularnewline
\hline
\end{longtable}

@@

\begin{lyxcode}

\end{lyxcode}
\bibliography{grid}
\end{document}

>>>>

\begin{tabular}{|>{\centering}p{0.15\columnwidth}|>{\centering}p{0.15\columnwidth}|c|>{\raggedright}p{0.4\columnwidth}|}
\hline 
Applicable BES State&
ARC Sub-state&
A-REX internal state&
Description\tabularnewline
\hline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{2}{2cm}{Pending}&
Accepting&
ACCEPTED&
Job is in the process of being submitted. This state is not recognised
by the A-REX yet. \emph{Accepted} is first reported state.\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Accepted&
ACCEPTED&
Job was submitted\tabularnewline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{7}{2cm}{Running}&
Preparing&
PREPARING&
Stage-in process is going on\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Prepared&
PREPARING+PENDING&
Stage-in process has finished\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Submiting&
SUBMIT&
Communication with local batch system is in process\tabularnewline
\hline 
&
Queued&
INLRMS&
Job entered local batch system but is not runnning now. This state
is not recognised by the A-REX yet. \emph{Executing} is reported instead \tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Executing&
INLRMS&
Job is being executed in local batch system\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Executed&
INLRMS, INLRMS+PENDING&
Job execution in local batch system has finished. The A-REX dos not
detect job states inside local batch system yet. As result this state
is reported only if job is \emph{Pending}.\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Killing&
CANCELING&
Communication with local batch system to terminate execution is in
process\tabularnewline
\cline{2-2} \cline{3-3} \cline{4-4} 
\multicolumn{1}{|c|}{}&
Finishing&
FINISHING&
Stage-out process is going on\tabularnewline
\hline 
Cancelled&
Killed&
FINISHED&
Job was stopped by explicit user request. The A-REX currently does
not remember this request. \emph{Failed} is reported instead.\tabularnewline
\hline 
Failed&
Failed&
FINISHED&
\multicolumn{1}{l|}{There was a failure during execution}\tabularnewline
\hline 
Finished&
Finished&
FINISHED&
Job finished successfully\tabularnewline
\hline 
Finished&
Deleted&
DELETED&
Job finished and was left in A-REX too long\tabularnewline
\hline 
All&
Pending&
PENDING&
Job is prevented from going to the next state due to some internal
limits; this sub-state appears in parallel with other sub-states\tabularnewline
\hline 
All&
Held&
&
Job processing is suspended on client request; this sub-state appears
in parallel with other sub-states. This state is reserved for future
and is not implemented yet.\tabularnewline
\hline
\end{tabular}
\end{table}

@@

\begin{itemize}
\item \emph{gm-kick} - Sends a signal to the A-REX though a FIFO file to
wake it up. It's used to increase responsiveness of A-REX.
\item \emph{CEinfo.pl} - Collects and generates information about computing
resource as XML document in NorduGrid and Glue 2 format.
\end{itemize}
The following modules are always run under the Unix account to which
a Grid user is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Currently supported LRMS are PBS/Torque, Condor, LoadLeveler, LSF
and SGE. Also \emph{fork} pseudo-LRMS is supported for testing purposes.
This module is responsible for job submission to the LRMS.
\item \textit{cancel-{*}-job} - This script is for canceling jobs which
have been already submitted to the LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the A-REX about completion of jobs. It's implementation for PBS uses
server logs to extract information about jobs. If logs are not available
it uses the less reliable \emph{qstat} command for that. Other backends
use different techniques.
\end{itemize}
\begin{comment}In addition, there is also an administration utility:

@@

\begin{itemize}
\item \textit{session root directory} - This is the directory in which a
user's SDs are created. It's location is configurable per UNIX user.
Several (or even all) users may share the same session root directory.\\
The A-REX need to have permission to create new files and directories
in the session root directory. If A-REX is run under a dedicated user
account, that account needs full permissions in the \textit{session
root directory}.\\
If A-REX is run under the \emph{root} account, make sure \textit{session
root directory} resides on a file system which does not limit the
capabilities of the \emph{root} user (as does for example NFS with
\emph{root\_squash} option).\\
If there is a need to run A-REX under the \emph{root} account (to
be able to run jobs in LRMS under different users' accounts, for example)
but there is no way to provide a suitable \textit{session root directory,}
use the \emph{norootpower} command in configuration file. In that
case A-REX will use the identity of the local user to which a Grid
identity is mapped to access the \textit{session root directory}.
Hence those users will need full access there.\\
The A-REX creates SDs with proper ownership and permissions for the
local identity used to run a job. Some file systems require users
to have \emph{execute} permission on the \textit{session root directory}
in order to access any file or subdirectory there.\\
In order for jobs to access their input files, session root directories
should be shared across cluster nodes. Otherwise, LRMS-specific methods
must be used to transfer files to execution nodes. 
\end{itemize}
\begin{comment} (LRMS\_COMMENT) For more information see section
\ref{sub:LRMS}.\end{comment}

\begin{itemize}
\item \textit{control directory} - In this directory A-REX stores an information
about accepted jobs. It must have full permissions there.\\
A subdirectory called \textit{log} is created there. It is used to
accumulate information about started and finished jobs. This information
is periodically sent to the desired \emph{logger service}(s). For
each job start and stop event, and for each logger service where that
event must be sent, a separate file is written. Once an event is sent,
the corresponding file is deleted.
\end{itemize}

@@

\begin{itemize}
\item \textbf{\textit{control}}\textit{=path username {[}username {[}...]]}
- This option initiates UNIX user as being serviced by the A-REX.
\textit{path} refers to the control directory (see section \ref{section: internals}
for the description of control directory). If the path is {*} the
default one is used - \$HOME/.jobstatus . \textit{username} stands
for UNIX name of the local user. Multiple names can be specified.
If the name is {*} it is substituted by all names found in file /etc/grid-security/grid-mapfile
(for the format of this file one should study the Globus project \cite{globus}).\\
Also the special name '.'(dot) can be used. Corresponding control
directory will be used for \textbf{any} user. This option should be
the last one in the configuration file. There is also command \textbf{\textit{controldir}}\textit{=path}.
It uses special username '.' and is always executed last independent
of placement in file.
\item \textbf{\textit{helper}}\textit{=username command {[}argument {[}argument
{[}...]]]} - associates external program with the local UNIX user.
This program will be kept running under account of the specified user.
\textit{username} stands for the name of the user. Special names can
be used: '{*}' - all names from /etc/grid-security/grid-mapfile, '.'
- root user. The user should be already configured with \textit{control}
option (except root, who is always configured). \textit{command} is
an executable and \textit{argument}s are passed as arguments to it.
\end{itemize}
\begin{comment}

Following are global commands specific to communication with underlying
LRMS (PBS in this case).

\begin{itemize}
\item \textbf{\textit{pbs\_bin\_path}}\textit{=path} - path to directory
which contains PBS commands.
\item \textbf{\textit{pbs\_log\_path}}\textit{=path} - path to directory
with PBS server's log files.
\end{itemize}
\end{comment}

\begin{itemize}
\item \textbf{\textit{gnu\_time}}\textit{=path} - path to \emph{time} utility.
\item \textbf{\textit{tmpdir}}\textit{=path} - path to directory for temporary
files.
\item \textbf{\textit{runtimedir}}\textit{=path} - path to directory which
contains \emph{runtimenvironment} scripts.
\item \textbf{\textit{shared\_filesystem}}\textit{=yes|no} - if computing
nodes have an access to session directory through a shared file system
like NFS. \begin{comment} Corresponds to an environment variable
RUNTIME\_NODE\_SEES\_FRONTEND (see section \ref{sub:LRMS}). \end{comment}
\item \textbf{\textit{nodename}}\textit{=command} - command to obtain hostname
of computing node.
\item \textbf{\textit{scratchdir}}\textit{=path} - path on computing node
where to move session directory before execution.
\item \textbf{\textit{shared\_scratch}}\textit{=path} - path on frontend
where \textbf{\textit{scratchdir}} can be found.
\end{itemize}
In the command arguments (paths, executables, ...) following substitutions
can be used:

@@

\begin{comment}(LRMS\_COMMENT)

@@

\end{comment}

@@

Please read \emph{User Guide} \cite{userguide1} for usage instructions.

@@

\medskip{}
Provided by uploader component (URL is replaced by destination of
output file) :\\
\begin{longtable}{|p{2.5in}|p{4in}|}
\hline 
\emph{Error string}&
\emph{Reason/description}\tabularnewline
\hline 
Internal error in uploader&
Generic error \tabularnewline
\hline 
Output file: URL - unknown error&
Generic error \tabularnewline
\hline 
Output file: URL - unexpected error&
Generic error \tabularnewline
\hline 
User requested to store output locally URL&
Destination is URL of type \emph{file}.\tabularnewline
\hline 
Output file: URL - bad source URL&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - bad destination URL&
Destination URL is either malformed or not supported\tabularnewline
\hline 
Output file: URL - failed to resolve source locations&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed to resolve destination locations&
Problems related to Data Indexing service.\tabularnewline
\hline 
Output file: URL - failed to register new destination file&
-//-\tabularnewline
\hline 
Output file: URL - can't start reading from source&
User request to store output file, but there is no such file or there
are problems accessing session directory\tabularnewline
\hline 
Output file: URL - can't start writing to destination&
Problems with Data Storing services\tabularnewline
\hline 
Output file: URL - can't read from source&
Problems accessing session directory\tabularnewline
\hline 
Output file: URL - can't write to destination&
Problems with Data Storing services\tabularnewline
\hline 
Output file: URL - data transfer was too slow&
Timeout during transfer\tabularnewline
\hline 
Output file: URL - failed while closing connection to source&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed while closing connection to destination&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed to register new location&
Problems related to Data Indexing service.\tabularnewline
\hline 
Output file: URL - can't use local cache&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - system error&
Operating System returned error code where unexpected\tabularnewline
\hline 
Output file: URL - delegated credentials expired&
Access to destination requires credententials and they are either
outdated or missing (not delegated).\tabularnewline
\hline
\end{longtable}

\begin{comment}(LRMS\_COMMENT)

@@

\end{comment}

\bibliography{grid}
\end{document}
